---
title: "Coding Finance Chap1 ~ Basic Financial Calculations"
author: "LordZhiHao"
date: '2022-05-30'
output: html_document
---

### Basic Financial Calculations

## #BUILD UP

Installing and loading of important libraries:

1. tidyQuant - [WEBSITE](https://cran.r-project.org/web/packages/tidyquant/index.html)
2. DT - [WEBSITE](https://rstudio.github.io/DT/)


``` {r, echo = T, results = 'hide', message = F}

# Install and load libraries

#install.packages("tidyquant")
#install.packages("DT")

library(tidyquant)
library(DT)
library(tidyverse)
library(dplyr)
library(magrittr)

options(scipen = 100, 
        digits = 4)

```

## #1 NET PRESENT VALUES 

Q1 :
 If we receive $100 per year for the next five years what is the net present value if the current interest rates are 5%. Let us set up our cash flow table.
 
```{r, echo = T, results = 'hide'}

# Create a interest rate variable 

r <- 0.05

# Create a cash flow table 

cfTable <- tribble(~Year, ~Cash_Flow,
                  2022, 100,
                  2023, 100,
                  2024, 100, 
                  2025, 100,
                  2026, 100)

cfTable %>% 
  datatable() %>% 
  formatRound(columns = 'Cash_Flow',
                        digits = 2)
```

# Calculating net present value 

```{r, echo = T, results = 'hide'}

# Calculating present value

cfTable %>% 
  mutate(present_value = Cash_Flow / (1 + r) ^ (Year - 2021)) %>% 
  datatable() %>% 
  formatRound(columns = 'present_value', 
              digits = 2)

cfTable

# net present value

npv <- cfTable %>% 
  mutate(present_value = Cash_Flow / (1 + r) ^ (Year - 2021)) %>% 
  select(present_value) %>%
  sum()

npv <- round(npv, 2)

```

Net present value is `r npv`. 

# Calculating present value of finite annuity

In some cases we have to calculate the present value of stream of equal payments. Many times a lottery winner is given the option to get the money in equal payments or a lump sum. Let us suppose you win a $50000 lottery and you are given two options

Receive five equal payments of $10000
Lump sum $45000.
Let us suppose the bank pays 5% interest rate on the deposit.

At first glance it may appear that five equal payments of $10000 is better than $45000. After all $50000 is more than $45000. To solve this problem we can calculate the net present value of those payments as below.

```{r, echo = T, results='hide'}

# creating the table and variables

pmt <- 10000
r <- 0.05

cfTable <- tibble(Year = 2022:2026,
                  pmt = pmt)
cfTable 

npv <- cfTable %>%
  mutate(present_value = pmt / (1 + r) ^ (Year - 2021)) %>% 
  select(present_value) %>%
  sum()

```

The NPV of the annuity is `r round(npv, 2)`. 


